The Sun lithium problem and how magnetic braking could help to solve it

### 2019-11-17 project/overleaf/paper.tex
Paper v3.3.1
Después de reunión 15.11.19

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

@@

@@

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Rotation, magnetic braking \& Li abundances]{Effects of rotation and magnetic braking on Li abundances for solar-type stars}

@@

% Abstract of the paper
\begin{abstract}
\textit{Goals.} The study of lithium (Li) surface abundance in the Sun and the influence of rotation and magnetic braking (MB) on its depletion during pre-main sequence (PMS) and main sequence (MS).
\newline\textit{Methods.} The effects of rotational mixing and of the rotational hydrostatic effects on Li abundances are studied by simulating several grids of PMS \& MS rotating and non-rotating models. The previous effects are combined with the additional impact of the MB (with magnetic field intensities ranging between 3.5 and 5.5 G) and the data obtained from the two set of simulations are confronted by comparing a number of parameters: Li abundances, angular velocity, effective temperature and size of convective zone.
\newline\textit{Results.} The surface Li abundance for the Sun like models at the end of the PMS and throughout the MS is decreased when rotational effects are included, i.e. the Li depletion rate for rotating models is higher than for non-rotating ones. However, this effect is attenuated when the braking effect produced by a magnetic field is present. This physical phenomena, generally neglected in star evolutionary models, impacts also the effective temperature ($T_{eff}$) of the star and therefore its location in the HR diagram. The impact of MB in Li depletion is sensitive to the magnetic field intensity: the higher it is, the lower the Li destruction. A direct link between the magnetic fields and the convective zone (CZ) size is observed: stronger magnetic fields produce shallower CZ's. Another important result of this study is that MB effect must be taken into consideration during PMS if we aim to reproduce Li abundances in young clusters.
\end{abstract}

@@

The denominated standard models, models that include convection only as a mixing process and do not consider other transport options like diffusion or angular momentum loss (AML), have been mainly involved in the elaboration of these predictions \citep{Sestito2005}. Li is destroyed at a temperature $T_{Li} \approx 2.5 x 10^6\; K$ when a Li atom collides with a proton producing two He atoms, something that takes place during proton-proton type II reactions (P-P II), and is therefore directly destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches that value. The Sun in particular and the solar-type stars in general are characterized by having a CZ that covers much of the stellar radius during the PMS which causes that their lower limit to exceed $T_{Li}$ \citep{Iben1965}. This depletion stops during the approach to the zero-age MS (ZAMS) when the convection zone retreats and the temperature at BCZ is cooler than $T_{Li}$. In the standard stellar models only the mass and the initial chemical composition determine at what distance from the stellar surface $T_{Li}$ temperature is reached, therefore stars within clusters with similar mass are expected to reach the ZAMS with equal surface Li abundances. Furthermore, they should also show a very similar Li evolution until their approach to the terminal-age MS (TAMS). During this same period the convection mechanisms unleash a mixing process that homogenizes the chemical composition of the convective envelope, from its lower limit till the star surface. However, different abundances of Li have been observed for different star populations \citep[][and references therein]{Somers2014}.\par

Theoretical models of stars don't inform about the initial amount of Li that a star has, they only describe how it is exhausted. Therefore, in order to make an accurate estimate of the initial abundance of Li, it is a prerequisite to be able to compare observations and models beforehand. Our Sun represents a unique exception, since it allows us to know the current abundance of this element in its photosphere, $A(\isotope[7]{Li}) = 1.1 \pm 0.1 \, dex$ \citep{Jeffries2004}, where $A(\isotope[7]{Li})$ is defined according to eq. (\ref{eq:A_Li})\par

@@

On the other hand, the initial abundance of $A(\isotope[7]{Li}) = 3.34 \, dex$ is obtained from meteorite measurements \citep{Randich2006}. According to the theory for newly born stars, we have that the initial abundance of Li can be estimated fairly accurately from photospherical measurements on T-Tauri type stars, or from the hottest F stars forming part of slightly older clusters. For the latter, the current theory of stellar evolution suggests that Li must not yet be exhausted. The results obtained from the measurements on both types of stars allow us to fix the initial abundance of Li in the interval $3.0 \, dex < A(\isotope[7]{Li}) < 3.4 \, dex$ \citep{Randich2006}.\par

The impact of rotation both on PMS and Li depletion for solar-type stars has already extensively debated in the past \citep{Pinsonneault1997,Jeffries2004,Somers2014} and revised more recently on the basis of the availability of more accurate measures \citep{Bouvier2016}. However, these previous studies focused mainly on the hydrostatic effects and didn't consider the influence of a coupling between the star magnetic field and its possible spin-down effect. It's thought that AML has a direct influence in the mixing processes and that transport of the momentum could be produced by a number of relevant mechanisms: mass loss, magnetic fields and gravity waves (g-modes). Concerning the two latter, gravity waves \citep{Charbonnel2005} and magnetic fields \citep{Eggenberger2009} have the property of transmiting angular momentum (AM) much more effectively than inducing mixing \citep{Denissenkov2007}. As a consequence of this increment in the AM transport efficiency, the amount of differential rotation between the radiative and the convective zones of the star is reduced (a solid-body rotation is fostered), as are the induced rotational instabilities. Of particular interest for this work is the impact caused by the magnetic field of the star in the AML, more specifically the effect produced on the solar wind particles that are forced to rotate in sync with it. As a consequence of this coupling a torque is induced which slows down the star, i.e. the magnetic braking effect. These magnetic fields would originate in the limits between the radiative and convective zones of the star, in the so-called tachoclines, due to the rotational difference between both zones \citep{Guerrero2016}. \par

Today, standard models are not able to replicate in a satisfactory way the observed values of Li abundance on star surfaces with the predictions that these models yield. This lack of agreement makes us think that certain physical mechanisms that influence the destruction of Li are either being modelled improperly or simply not being taken into account, e.g. the magnetic braking. In view of these findings, it seems evident that there are still open questions to be resolved on the subject of the processes that participate, directly or indirectly, in the mixing mechanisms that are not incorporated in the standard stellar models.\par

Therefore, it's essential a proper consideration of the interactions between rotation and magnetic fields when it comes to study the AM distribution. We have adapted the one-dimensional stellar evolution code used in this work to model MB and included its effect on the equations which govern both AM distribution and stellar structure. This approach produces a more complete, though still reasonably simple, model for AM evolution. It's worth to mention that in this study we don't only focused in investigating the effects of rotation on the Li content during PMS and MS but also on the indirect role played by the MB on the Li destruction as a result of its influence on the rotational history of solar-type stars.\par

According to the prevalent theory of rotational mixing, southern circulation and internal shears leads to a homogenization of the material depleted of Li found within the CZ, causing that the inner layers where there is a temperature higher than $T_{Li}$ and where Li is destroyed, been mixed with those cooler above them. These mechanism causes a positive feedback loop in which the cooler material sinks, is heated again and the remaining Li is destroyed, totally or partially, before proceeding to rise once more into cooler layers. \par

In the remainder of this paper, we describe a physically sound yet computationally simple model extension to Modules for Experiments in Stellar Astrophysics star evolution code (MESA; \citeauthor{Paxton2011} \citeyear{Paxton2011}, \citeyear{Paxton2013}, \citeyear{Paxton2015}, \citeyear{Paxton2018}, \citeyear{Paxton2019}) for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar winds. In its simplest form, the implementation of the model requires the specification of both the surface magnetic field strength ($B$) and the initial rotation rate ($\Omega$) at star surface. The models include already rotation during the PMS as there evidences that advocate for a strong established relationship between Li destruction and rotation on that phase (\citeauthor{Bouvier2016} \citeyear{Bouvier2016}, \citeyear{Bouvier2018}), and the numerical simulations trace the rotational history and A(Li) of a $1\, M_{\sun}$ star for a variety of initial values for $B$ and $\Omega$.\par

\subsection{Effects of magnetic fields}
The origin of magnetic fields is still unknown. The organized magnetic fields detected in some O stars \citep{Wade2010} might be fossil fields \citep[see][for details about the theory of fossil magnetic field]{Dudorov2014}, or fields produced through a dynamo mechanism \citep{Cantiello2009}. Besides that, the presence of magnetic fields with intensities of the order of kG \citep{Hussain2014} has been a sine qua non condition for dealing with certain data observed in T Tauri type stars in young, accreting PMS systems \citep{Johns-Krull2007}.\par

The mass of stars plays a determining role in the existence or not of an external convective envelope. In turn, the presence of this convective region conditions the evolution of magnetic fields. Massive stars do not have external convective zones although they can expose a convective core. On the other hand, the less massive stars such as the Sun, have convective covers and a radiative core. In both types of stars strong magnetic fields can be generated in the tachocline but with notable differences (see \citep{Chabrier2006} \citet{Charbonneau2010}\ for details).\par

\subsubsection{Internal Magnetic Fields}
Our Sun is an active star as far as the magnetic field is concerned. According to the theoretical models this magnetic field is generated by a dynamo type effect in the tachocline, the intermediate zone between the radiative core and the CZ \citep{Aschwanden2014}, and has an intensity of approximately $B\approx10^5\, G$. Additionally, these models predict that because of the differential rotation existing between the core and CZ, the lines of the magnetic field that at the moment of their generation have a poloidal orientation will be affected by the differential rotation and forced to adopt a toroidal typology (the $\omega$-effect). On the other hand, we have that the opposite case can also occur, that is to say, starting from a toroidal magnetic field to obtain a poloidal configuration. This would be possible on the one hand, thanks to the convective movements of material that are produced in the CZ of the star and on the other, to assume that this material behaves as a perfect conductor. This second condition allows to assume that the magnetic field lines are "frozen" in the conductive fluid and therefore must move in solidarity with it ($\alpha$-effect).\par

\subsubsection{Surface Magnetic Fields} \label{surf_mf}
The intensity of the magnetic field measured on the Sun's surface is about $B\approx1\, G$, twice the value of the Earth's one. Additionally its topology is varied and complex. The dynamo model establishes that the magnetic field produced inside the star reaches the surface with the help of convective processes taken place in the CZ and gives rise to magnetically active regions on they reach the solar surface. Besides, rotating stars that expose a significant outer CZ can produce surface magnetic fields through a dynamo mechanism \citep[e.g.][]{Brandenburg2004,Charbonneau2010,Brun2017}. Whatever the origins of surface magnetic fields are, it's expected to couple to the wind mass-loss and, if this is strong enough, to produce MB \citep[e.g.][]{UdDoula2002,Ud-Doula2007,Ud-Doula2008,Meynet2010}.\par

\subsection{Effects of magnetic braking}
In its most widely accepted approach, the MB is linked to Skumanich's work in which he develops an empirical law of evolution of the MB according to the age of the star. To calibrate this law, Skumanich relied on data obtained from type G stars found in the MS \citep{Skumanich1972}. The influence of the MB on the evolution of the star is closely related to the transport of the AM and by extension, influenced by how the chemical elements are transported within it. According to this approach \citep{Meynet2010}, two main mechanisms are distinguished attending to how the transport of the AM is produced which influence the mixing of these elements:

\begin{enumerate}
    \item Differential rotation: the AM transport is pushed by southern currents and shear instabilities.
    \item Solid body rotation: when the transport of AM is very efficient, the solid body rotation is the prevailing mechanism maintained throughout the entire MS phase and the chemical elements are transported by the southern currents.
\end{enumerate}

The AML caused by MB depends directly on the amount of mass lost by the star due to the stellar winds. The estimated mass loss ratio for a solar-type star, approximately $10^{-13}M_{\odot} \; yr^{-1}$, and the resulting AML derived from that value (refer to \ref{mod_mb} for details) can be considered relatively modest as to decisively influence the evolution of the star. Furthermore, the combined effect of convective movements and differential rotation lead to the generation of magnetic fields. Those magnetic fields end up reaching the surface of the star \citep{Langer2012}, so that the AML can be increased in several orders of magnitude. That increment is as a consequence of the magnetic field that forces the stellar wind ionized particles to turn with the same angular speed, winds which extends a distance several times the radius of the star \citep[see][for more details]{UdDoula2002,Ud-Doula2007,Ud-Doula2008}.

This paper is structured as follows: Section \ref{sec_2} gives an overview of rotation and magnetic fields physical processes and how do they impact on the star evolution. We also summarized here the most relevant parameters configured in MESA for the model simulations. In Section \ref{sec_3}, the model outputs are presented and discussed in detail. Finally, Section \ref{sec_4} finishes this paper with a discussion about open points and future works.\par

\section{Method} \label{sec_2}
\begin{table}
	\centering
	\caption{Summary of adopted physics in MESA \citep[based on][]{Choi2016}.}
	\label{tab:phy_mesa}
	\begin{tabular}{ll} 
		\hline
		Parameter & Adopted prescriptions and values\\
		\hline
		Solar Abundance & $X_{\odot}=0.7154, Y_{\odot}=0.2703, Z_{\odot}=0.0142$\\
		Equation of State & OPAL+SCVH+MacDonald+HELM+PC\\
		Opacity & OPAL Type I for log T $\geq$ 4 \\ & Ferguson for logT $\leq$ 4\\
		Reaction Rates & JINA REACLIB\\
		Boundary Conditions & ATLAS12; $\tau$=100 tables + photosphere\\
		Difussion & Track \isotope[1]{H}, \isotope[2]{He}, \isotope[7]{Li}, \isotope[7]{Be}\\
		Rotation & Differential rotation at PMS \& MS\\
		Convection & MLT + Ledoux, $\alpha_{MLT}$ = 1.82\\
		Overshoot & time-dependent, diffusive, \\ & $f_{ov,core}=0.0160$,\\ 
		& $f_{ov,sh}=0.0174$\\
		Semiconvection & $\alpha_{sc}=0.1$\\
		Thermohaline & $\alpha_{th}=666$\\
		Rotational Mixing & Include SH, ES, GSF, SSI \& DSI\\
		Magnetic Effects & Magnetic braking based on idealized \\ & monopole field\\
		Magnetic Field & B(G) variable between [3.5 - 5.5]\\
		Mass Loss & activated, $\Dot{M}_{max} = 10^{-3} \: M_{\sun} \: yr^{-1}$\\
		Angular Moment Loss & activated, $\Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A}$\\
		\hline
	\end{tabular}
\end{table}

We define here also some conventions and assumptions adopted in the paper. For the models presented in this paper we adopt solar-scaled abundances and assumes the \citet{Asplund2009} protosolar birth cloud metallicity ($Z = Z_{\odot, proto} = 0.0142$) not the current photospheric metallicity ($Z = Z_{\odot, photosphere} = 0.0134$) as the reference value. We also adopt the following nominal values to express stellar properties values in SI units $R_{\odot} = 6.957x10^{10}\, cm$ and $M_{\odot} = 1.988x10^{33}\, g$ which are consistent with IAU resolutions \citep{Mamajek2015}. For a detailed description of the physics adopted in this paper, refer to \citet{Choi2016}. That work, in particular its section 4.1, has been used as a starting point for ours regarding the parameterization of the MESA project which is calibrated to reproduce the measured element abundances on the solar surface.\par

In this section we review the core physic aspects simulated in MESA and the most relevant parameters adopted in the numerical simulations. In its simplest version, by activating rotation in the model, the evolutionary computational code MESA takes into account the following effects:
%\vspace{-1pc}
\begin{enumerate}
    \item AM transport from the radiative interior to the convective envelope in response to the rotational deceleration of the star surface layers\label{itm:1}.
    \item AM redistribution associated with changes in internal structure during the process of contraction to the MS\label{itm:2}.
    \item AML as a result of the torque applied to the convection zone by a magnetically coupled wind\label{itm:3}.
\end{enumerate}

The enumerated effects \ref{itm:1} and \ref{itm:2} are standard features offered by MESA. On the contrary, the effect \ref{itm:3} isn't available as an out-of-the-box feature during the elaboration of this paper and represents our extension to MESA.

The rest of this section is devoted to describe the most relevant aspects and assumptions made in modelling the evolution of rotation, magnetic braking and angular momentum in MESA.\par

\subsection{Modelling rotation}
The rotation to which a star is subjected has profound implications for its inner structure and, therefore, for the way in which it evolves throughout its life. It is also responsible for inducing through hydrodynamic instabilities the mixing of the chemical elements that make it up. In order to be able to correctly model those induces mixing effects consistently with the observational data, it is necessary to analyze and understand more precisely the mechanisms that govern the distribution of AM \citep{Pinsonneault1997,Maeder2000}, since in this area the theoretical models also offer deficiencies to explain the observations \citep{Denissenkov2007}. \par

Stars have a three-dimensional structure and the equations that model their evolution must take this fact into account. However, it has been proven sufficient to solve the equations in only one dimension, extrapolating the results to the other two (shellular approximation \citep{Meynet1997}) MESA uses this simplification to accelerate computational calculations \citet{Paxton2013}.\par

\subsection{Modelling magnetic braking} \label{mod_mb}
In this section, we describe a physically sound yet computationally simple model extension to Modules for Experiments in Stellar Astrophysics star evolution code (MESA; \citeauthor{Paxton2011} \citeyear{Paxton2011}, \citeyear{Paxton2013}, \citeyear{Paxton2015}, \citeyear{Paxton2018}, \citeyear{Paxton2019}) for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar winds. In its simplest form, the implementation of the model requires the specification of both the surface magnetic field strength ($B$) and the initial rotation rate ($\Omega$) at star surface. The models include already rotation during the PMS as there evidences that advocate for a strong established relationship between Li destruction and rotation on that phase (\citeauthor{Bouvier2016} \citeyear{Bouvier2016}, \citeyear{Bouvier2018}), and the numerical simulations trace the rotational history and A(Li) of a $1\, M_{\sun}$ star for a variety of initial values for $B$ and $\Omega$.\par

In the presence of mass-loss, AM is removed by the material in the stellar wind, resulting in a spin-down effect. This is quite important for example in massive stars, that are known to be rapidly rotating and have strong stellar winds. It has been also observed that strongly magnetic intermediate-mass stars typically have rotation rates much slower than other stars in their parent population \citep{Mathys2006}. In those stars, the presence of such magnetic field will interact with the mass loss and the Alfv\'{e}n radius ($R_{A}$) plays an important role in this interaction. The $R_{A}$ is defined as the point in which the magnetic field energy density $B^{2}/8\upi$, marked by magnetic field strength ($B$), and the kinetic energy density $\frac{1}{2}\rho\nu^{2}$, defined by the mass density ($\rho$) and velocity of the stellar wind $(\nu)$, are balanced. That is, where the coefficient between both is equal to 1. In case that $R_{A}$ is greater than the star radius, then the wind flow will have to follow the magnetic field. As a consequence, the material will leave the stellar surface with a higher specific AM, as the co-rotation radius has increased and it will roughly correspond to the $R_{A}$. This effect is known as MB.\par

In the present work we do not take into account either the influence of internal magnetic fields or their existence during the T-Tauri phase. We adopt a simple and pragmatic theoretical approach to establish their occurrence when the star is reaching the ZAMS phase. The adopted criterion is based on the simultaneous existence of an extensive convective layer and a radiative core. From this moment on, the MB routine will be activated, acting as an additional mechanism to those existing in the MESA evolutionary code that participates in the star AML. Additionally, we assume that the magnetic field does not vary its intensity throughout the evolution of the star until it reaches the TAMS.\par

The energy ratio (eq. \ref{eq:wind_conf}) is denominated wind confinement magnetic parameter ($\eta_*$) and defines a characteristic parameter for the relative effectiveness of the magnetic fields in circumscribing and/or channeling the wind outflow \citep{UdDoula2002}.\par

@@

@@

In turn, of the observable characteristics of the solar wind, we especially care about the parameters related to the amount of mass that is able to pull out of the outer layers of the star in a given time interval, and the speed that reaches the wind itself measured at a great distance from the star, the terminal velocity ($\nu_\infty$). In general, a magnetically channeled outflow will have a complex stream geometry but for convenience, the expression (\ref{eq:mass_loss}) simply characterizes the wind strength in terms of a spherically symmetric mass-loss rate. $\nu$ can be characterized by the radial variation of outflow velocity in terms of the velocity law, where $r$ represents the distance from the center of the star to the point where the velocity is to be measured and $R_*$ is the radius of the star:
\begin{ceqn}
\begin{equation}
    \nu(r) = \nu_\infty (1-R_*/r)
\end{equation}
\end{ceqn}
where $\nu_\infty$ is the terminal wind velocity defined as the velocity that the wind, the outflowing matter, reaches at large distance from the central star, where it is not accelerated anymore by the wind driving force but its deceleration due to interaction with the interstellar medium (ISM) is negligible \citep{Niedzielski2002}.\par

The line-driven winds of massive OB stars have terminal velocities that scale with the photospheric escape velocity \citep{Lamers2000} according to:
\begin{ceqn}
\begin{align}
\nu_{esc} &= \sqrt{\frac{2GM_*}{R_*}} \label{eq:vesc} \\
\nu_\infty &\simeq 1.92 \;\nu_{esc}\\
\nu_\infty &\simeq 1.92 \; x \; 618 \; \Bigg(\sqrt{\frac{R_{\sun}}{R_*}\frac{M_*}{M_{\sun}}} \;\Bigg) \label{eq:vinf}
\end{align}
\end{ceqn}
where (\ref{eq:vesc}) is the Newtonian escape velocity from the stellar surface, $G$ is the gravitational constant and expression (\ref{eq:vinf}) represents the terminal velocity in terms of escape velocity, radius and mass of the Sun.\par

>>>>

% MNRAS is set in Times font. If you do not have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

@@

\input definitions.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Rotation, magnetic braking \& Li abundances]{The Sun lithium problem and how magnetic braking could help to solve it}

@@

% Abstract of the paper
\begin{abstract}
\textit{Goals.} The study of lithium (Li) surface abundance in the Sun and young stellar globular clusters which are seemingly anomalous in present-day scenarios, as well as the influence of rotation and magnetic braking (MB) on its depletion during pre-main sequence (PMS) and main sequence (MS).
\newline\textit{Methods.} The effects of rotational mixing and of the rotational hydrostatic effects on Li abundances are studied by simulating several grids of PMS \& MS rotating and non-rotating models. The previous effects are combined with the additional impact of the MB (with magnetic field intensities ranging between 3.0 and 5.0 G) and the data obtained from the two sets of simulations are confronted by comparing a number of parameters: Li abundances, angular velocity, effective temperature and size of convective zone.
\newline\textit{Results.} The surface Li abundance for the Sun like models at the end of the PMS and throughout the MS is decreased when rotational effects are included, i.e. the Li depletion rate for rotating models is higher than for non-rotating ones. However, this effect is attenuated when the braking effect produced by a magnetic field is present. This physical phenomena, generally neglected in stellar evolutionary models, impacts also the effective temperature ($\teff$) of the star and therefore its location in the HR diagram. The impact of MB in Li depletion is sensitive to the magnetic field intensity: the higher it is, the lower the Li destruction. A direct link between the magnetic fields and the convective zone (CZ) size is observed: stronger magnetic fields produce shallower CZ's. Another important result of this study is that MB effect must be taken into consideration during PMS if we aim to reproduce Li abundances in young clusters.
\end{abstract}

@@

The denominated standard models, models that include convection only as a mixing process and do not consider other transport options like diffusion or angular momentum loss (AML), have been mainly involved in the elaboration of these predictions \citep{Sestito2005}. Li is destroyed at a temperature $\tli \approx 2.5 x 10^6\; K$ when a Li atom collides with a proton producing two He atoms, something that takes place during proton-proton type II reactions (P-P II), and is therefore directly destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches that value. The Sun in particular and the solar-type stars in general are characterized by having a CZ that covers much of the stellar radius during the PMS which causes that their lower limit to exceed $\tli$ \citep{Iben1965}. This depletion stops during the approach to the zero-age MS (ZAMS) when the convection zone retreats and the temperature at BCZ is cooler than $\tli$. In the standard stellar models only the mass and the initial chemical composition determine at what distance from the stellar surface $\tli$ temperature is reached, therefore stars within clusters with similar mass are expected to reach the ZAMS with equal surface Li abundances. Furthermore, they should also show a very similar Li evolution until their approach to the terminal-age MS (TAMS). During this same period the convection mechanisms unleash a mixing process that homogenizes the chemical composition of the convective envelope, from its lower limit till the star surface. However, different abundances of Li have been observed for different star populations \citep[][and references therein]{Somers2014}.\par

Theoretical models of stars do not inform about the initial amount of Li that a star has, they only describe how it is exhausted. Therefore, in order to make an accurate estimate of the initial abundance of Li, it is a prerequisite to be able to compare observations and models beforehand. Our Sun represents a unique exception, since it allows us to know the current abundance of this element in its photosphere, $A(\isotope[7]{Li}) = 1.1 \pm 0.1 \, ex$ \citep{Jeffries2004}, where $A(\isotope[7]{Li})$ is defined according to Eq.~\ref{eq:A_Li}\par

@@

On the other hand, the initial abundance of $A(\isotope[7]{Li}) = 3.34 \, dex$ is obtained from meteorite measurements \citep{Randich2006}. According to the theory for newly born stars, we have that the initial abundance of Li can be estimated fairly accurately from photospheric measurements on T-Tauri type stars, or from the hottest F stars forming part of slightly older clusters. For the latter, the current theory of stellar evolution suggests that Li must not yet be exhausted. The results obtained from the measurements on both types of stars allow us to fix the initial abundance of Li in the interval $3.0 \, dex < A(\isotope[7]{Li}) < 3.4 \, dex$ \citep{Randich2006}.\par

The impact of rotation both on PMS and Li depletion for solar-type stars has already extensively debated in the past \citep{Pinsonneault1997,Jeffries2004,Somers2014} and revised more recently on the basis of the availability of more accurate measures \citep{Bouvier2016}. However, these previous studies focused mainly on the hydrostatic effects and did not consider the influence of a coupling between the star magnetic field and its possible spin-down effect. It is thought that AML has a direct influence in the mixing processes and that transport of the momentum could be produced by a number of relevant mechanisms: mass loss, magnetic fields and gravity waves (g-modes). Concerning the two latter, gravity waves \citep{Charbonnel2005} and magnetic fields \citep{Eggenberger2009} have the property of transmitting angular momentum (AM) much more effectively than inducing mixing \citep{Denissenkov2007}. As a consequence of this increment in the AM transport efficiency, the amount of differential rotation between the radiative and the convective zones of the star is reduced (a solid-body rotation is fostered), as are the induced rotational instabilities. Additionally, magnetic fields would also originate in the limits between these zones, in the so-called tachoclines \citep{Aschwanden2014, Guerrero2016}, which interact with the solar wind particles that are forced to rotate in sync with it and therefore slow down the star, i.e. the magnetic braking (MB) effect. \par

Today, standard models are not able to replicate in a satisfactory way the observed values of Li abundance on star surfaces with the predictions that these models yield. This lack of agreement makes us think that certain physical mechanisms that influence the destruction of Li are either being modelled improperly or simply not being taken into account, e.g. the magnetic braking. In view of these findings, it seems evident that there are still open questions to be resolved on the subject of the processes that participate, directly or indirectly, in the mixing mechanisms that are not incorporated in the standard stellar models. Therefore, it is essential a proper consideration of the interactions between rotation and magnetic fields when it comes to study the AM distribution. \par

\subsection{Effects of magnetic fields}
The origin of magnetic fields in the stars is still unknown. The organized magnetic fields detected in some O stars \citep{Wade2010} might be fossil fields \citep[see][for details ]{Dudorov2014}, or fields produced through a dynamo mechanism \citep{Cantiello2009}. Besides that, the presence of magnetic fields with intensities of the order of kG \citep{Hussain2014} has been a sine qua non condition for dealing with certain data observed in T Tauri type stars in young, accreting PMS systems \citep{Johns-Krull2007}.\par

The mass of stars plays a determining role in the existence or not of an external convective envelope. In turn, the presence of this convective region conditions the evolution of magnetic fields. Massive stars do not have external convective zones although they can expose a convective core. On the other hand the less massive stars, such as the Sun, have convective covers and a radiative core. In both types of stars strong magnetic fields can be generated in the tachocline but with notable differences \citep[e.g.][for details]{Chabrier2006,Charbonneau2010}.\par

\subsubsection{Internal Magnetic Fields}
Our Sun is an active star as far as the magnetic field is concerned. According to the theoretical models this magnetic field is generated by a dynamo in the tachocline and with an intensity about $B\approx10^{5}\,\Gauss$ \citep{Maeder2003a,Dudorov2014}. At the moment of their generation, the lines of that magnetic field will have a poloidal orientation and will be affected by the differential rotation and forced to adopt a toroidal typology (the $\omega$-effect). On the other hand, the opposite case can also occur, starting from a toroidal magnetic field to obtain a poloidal configuration. This would be possible thanks to the convective movements of material that are produced in the CZ of the star and to assume that this material behaves as a perfect conductor. This second condition allows to assume that the magnetic field lines are "frozen" in the conductive fluid and therefore must move in solidarity with it ($\alpha$-effect).\par

\subsubsection{Surface Magnetic Fields} \label{surf_mf}
The intensity of the magnetic field measured on the Sun's surface is about $B\approx1\,\Gauss$ \citep{Weber1967,DAntona2000,Morin2012}, twice the value of the Earth's one. Additionally its topology is varied and complex. The dynamo model establishes that the magnetic field produced inside the star reaches the surface with the help of convective processes taken place in the CZ and gives rise to magnetically active regions on the solar surface. Besides, rotating stars that expose a significant outer CZ can produce surface magnetic fields through a dynamo mechanism \citep[e.g.][]{Brandenburg2004,Charbonneau2010,Brun2017}. Whatever the origins of surface magnetic fields are, it is expected to couple to the wind mass-loss and, if this is strong enough, to produce MB \citep[e.g.][]{UdDoula2002,Ud-Doula2007,Ud-Doula2008,Meynet2010}.\par

\subsection{Effects of magnetic braking}
In its most widely accepted approach, the MB is linked to Skumanich's work \citep{Skumanich1972} in which he develops an empirical law of its evolution in parallel to the age of the star. To calibrate this law, Skumanich relied on data obtained from type G stars found in the MS. The influence of the MB on the evolution of the star is closely related to the transport of the AM. According to \citet{Meynet2010}, two main mechanisms are distinguished attending to how the transport of the AM is produced:

\begin{enumerate}
    \item Differential rotation: the AM transport is pushed by southern currents and shear instabilities.
    \item Solid body rotation: when the transport of AM is very efficient, solid body rotation is 
    maintained throughout the entire MS phase.
\end{enumerate}

The AML caused by MB depends directly on the amount of mass lost by the star due to the stellar winds. The estimated mass loss ratio for a solar-type star, approximately $10^{-14}\msun \; yr^{-1}$ \citep{Noerdlinger2008}, and the resulting AML derived from that value can be considered relatively modest as to decisively influence the evolution of the star. Furthermore, the combined effect of convective movements and differential rotation lead to the generation of magnetic fields. Those magnetic fields end up reaching the surface of the star, so that the AML can be increased in several orders of magnitude \citep{Langer2012}. That increment is a consequence of the magnetic field that forces the stellar wind ionized particles to turn with the same angular speed, winds which extends a distance several times the radius of the star \citep[see][for more details]{UdDoula2002,Ud-Doula2007,Ud-Doula2008}.

\section{Method} \label{sec_2}

In the remainder of this paper, we describe a physically sound yet computationally simple model which will be used as an extension to Modules for Experiments in Stellar Astrophysics stellar evolution code (MESA; \citeauthor{Paxton2011} \citeyear{Paxton2011}, \citeyear{Paxton2013}, \citeyear{Paxton2015}, \citeyear{Paxton2018}, \citeyear{Paxton2019}) for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar winds.

\subsection{Modelling magnetic braking} \label{mod_mb}
We'll start enumerating the most relevant aspects and assumptions made in modelling the evolution of rotation, magnetic braking and angular momentum in MESA for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar winds.\par

In the presence of mass-loss, AM is removed by the material in the stellar wind, resulting in a spin-down effect. This is quite important for example in massive stars, that are known to be rapidly rotating and have strong stellar winds. It has been also observed that strongly magnetic intermediate-mass stars typically have rotation rates much slower than other stars in their parent population \citep{Mathys2006}. In those stars, the presence of such magnetic field will interact with the mass loss and the Alfv\'{e}n radius ($R_{A}$) plays an important role in this interaction.\par

A relevant parameter to characterize the influence of a given magnetic field on the stellar wind is the denominated wind confinement magnetic parameter ($\eta_*$). It represents a energy ratio (Eq.~\ref{eq:wind_conf}) and defines a characteristic parameter for the relative effectiveness of the magnetic fields in circumscribing and/or channeling the wind outflow \citep{UdDoula2002}.\par

@@

Making using of ($\eta_*$), the $R_{A}$ is defined as the point in which the magnetic field energy density $B^{2}/8\upi$, marked by magnetic field strength ($B$), and the kinetic energy density $\frac{1}{2}\rho\nu^{2}$, defined by the mass density ($\rho$) and velocity of the stellar wind $(\nu)$, are balanced. That is, where the coefficient between both is equal to 1. In case that $R_{A}$ is greater than the star radius, then the wind flow will have to follow the magnetic field. As a consequence, the material will leave the stellar surface with a higher specific AM, as the co-rotation radius has increased and it will roughly correspond to the $R_{A}$. This effect is known as MB.\par

@@

In turn, of the observable characteristics of the solar wind, we especially care about the parameters related to the amount of mass that is able to pull out of the outer layers of the star in a given time interval, and the speed that reaches the wind itself measured at a great distance from the star, the terminal velocity ($\nu_\infty$). In general, a magnetically channeled outflow will have a complex stream geometry but for convenience, the Eq.~\ref{eq:mass_loss} simply characterizes the wind strength in terms of a spherically symmetric mass-loss rate. $\nu$ can be characterized by the radial variation of outflow velocity in terms of the velocity law, where $r$ represents the distance from the center of the star to the point where the velocity is to be measured, and $R_*$ is the radius of the star:
\begin{ceqn}
\begin{equation}
    \nu(r) = \nu_\infty (1-R_*/r)
\end{equation}
\end{ceqn}
where $\nu_\infty$ is the terminal wind velocity defined as the velocity that the wind, the outflowing matter, reaches at large distance from the central star, where it is not accelerated anymore by the wind driving force but its deceleration due to interaction with the interstellar medium (ISM) is negligible \citep{Niedzielski2002}.\par

The line-driven winds of massive OB stars have terminal velocities that scale with the photospheric escape velocity \citep{Lamers2000} according to:
\begin{ceqn}
\begin{align}
\nu_{esc} &= \sqrt{\frac{2GM_*}{R_*}} \label{eq:vesc} \\
\nu_\infty &\simeq 1.92 \;\nu_{esc}\\
\nu_\infty &\simeq 1.92 \; x \; 618 \; \Bigg(\sqrt{\frac{R_{\sun}}{R_*}\frac{M_*}{\msun}} \;\Bigg) \label{eq:vinf}
\end{align}
\end{ceqn}
where Eq.~\ref{eq:vesc} describes the Newtonian escape velocity from the stellar surface, $G$ is the gravitational constant and Eq.~\ref{eq:vinf} represents the terminal velocity in terms of escape velocity, radius and mass of the Sun.\par
